Large-Scale Marine Protected Areas in the World's Largest Tuna Fishery

### 2019-02-07 docs/PIPA_Palau.tex
Updated figures and MS

\documentclass[9pttwoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

@@

\begin{abstract}
The Parties to the Nauru Agreement (PNA) is a system for managing the world's largest skipjack tuna fishery, covering an area of ~14.6 million km\textsuperscript{2} in the Pacific Ocean. Countries within the PNA have recently started to prohibit fishing within large parts of the PNA area. It is still unclear what the effects of these prohibitions have been and what effects are likely to occur in the future. We use identification of fishing activity via
Automatic Identification Systems (AIS) and causal inference techniques
to provide the first analysis of behavioral changes and spatial
redistribution of tuna purse seiners due to the implementation of a
Large-Scale Marine Protected Area (LSMPA) in the Pacific Ocean. Our work
provides three main findings: 1) aggregate fishing effort remains
relatively unaffected; 2) vessels that fished inside the protected area
redistribute to adjacent waters (including the high seas); and 3) we observe a crowding effect for
the first months after implementation. This suggests that rights to fish in PNA waters might now be less valuable. Our results not only provide an
impact evaluation of the effect of LSMPAs on fishing activity, but
provide insights into vessel redistribution dynamics, which may have
ecological and economic implications for marine conservation and fisheries
management. As countries continue to implement LSMPAs as a way to
reach ocean protection targets, managers should
consider how fishing effort will change in space and through time to
ensure that fishing effort is not just displaced elsewhere, leading to
overfishing in adjacent waters. We do not observe an increase in daily fishing hours for vessels that used to fish within the LSMPA, compared to a control
group of vessels. Thus, for vessels that have chosen to continue to operate following the closure, we do not observe evidence that they are being forced to exert more effort as a result of the closure. Unfortunately, without detailed data on vessel catch, we cannot conclude whether vessels are better or worse off financially following the closure. We use our results to inform predictions of the impacts of a proposed LSMPA in Palau (a PNA member). Our estimates of the potential losses range from \$2.5 to \$11 million annually. This wide range hinges on what Palau's future PNA allocation will be and the extent to which US aid will continue through non-fishing channels. 
\end{abstract}

@@

%Next steps: I think we need to make the buffers around the EEZs 5 degrees instead of 1 degree (redistribution of effort section)
%Might change the specification of the redistribution regression
%
%(QUESTION FOR JC: WHEN IS THIS (selling of rVDS rights)? And wasn't the closure expected?).

@@

\begin{figure}
\centering
\includegraphics{img/LSMPAs_map.png}
\caption{\label{fig:LSMPAs_map}Large Scale Marine
Protected Areas. The map shows all areas larger than 250,000
Km\textsuperscript{2}. Areas in dark blue meet at least one of these
conditions: reported no take area is greater than 0, are categorized as
IUCN Ia or Ib, their designated English name is `Protected Area'.}
\end{figure}

@@

\hypertarget{nauru-agreement-and-the-phoenix-island-protected-area}{%
\subsection{Nauru agreement and the Phoenix Islands Protected
Area}\label{nauru-agreement-and-the-phoenix-island-protected-area}}

The Nauru Agreement was established in 1982 by a select group of Pacific
island nations to manage their important tuna resources more effectively. Parties to the Nauru Agreement (PNA) Members
include the Federated States of Micronesia, Kiribati, the Marshall Islands,
Nauru, Palau, Papua New Guinea, the Solomon Islands, and Tuvalu. Tokelau joined in PNA group in 2012 and started selling access rights in 2013. The Nauru
Agreement regulates access of foreign vessels (\emph{i.e.} those from
non-PNA countries). Holding \textasciitilde{}80\% of historical
skipjack tuna purse seine grounds within their Exclusive Economic Zones, PNA
countries have achieved greater bargaining power when providing fishing access to foreign
fleets \cite{havice_2010}.

The cooperation that emerged under the Nauru Agreement allowed for subsequent
agreements that strengthened fisheries management, like the Palau
Agreement, which limited the number of purse seiners at 205 vessels from
1995-2007.\footnote{See \cite{havice_2010} for a detailed description of
	the Nauru, Palau, and Federal States of Micronesia agreements.} However, the most notable regulation is
the approach to manage fishing effort: a Vessel Day Scheme (VDS)
implemented in 2007 \citep{havice_2013}. This effectively modified how
fishing effort was managed, from total number of vessels (under the Palau
Agreement) to total fishing hours. 
Under the purse seine VDS, a total number of annual vessel days for the entire fishery is agreed upon by the PNA parties. Each party’s allowable effort (PAE) is then calculated based on historic effort and biomass within each party’s EEZ. Sixty percent of the PAE is calculated based on EEZ effort over the last seven years and 40\% of the PAE is calculated based on the 10-year average of each country’s share of estimated skipjack and yellowfin biomass within its EEZ.\footnote{This is explained in more detail in Article 12.5 of the 2012 Amendment to the Palau Agreement and in \cite{Hagrannsoknir2014}}. A minimum benchmark fee is set for purse seine vessel days which each party can transfer (i.e., sell to another PNA member) or sell to the highest bidder (i.e., sell to a fishing company). Vessel days can be transferred to fish in any PNA party’s EEZ, without penalty to the transferred parties' allocations \cite{PNA2016}.\footnote{The process of transferring days between parties is described in Article 7 of the Management Scheme \cite{PNA2016}.} The mean value of a vessel day has steadily increased since 2007 \citep{havice_2013}. 
Although detailed records on sales or transfers of vessel days are not publicly available \citep{havice_2013,yeeting2018stabilising}, a minimum benchmark fee of \$8,000/day was set from 2015 \citep{PNA2014a}.

@@

\begin{figure}
\centering
\includegraphics{img/PNA_map.png}
\caption{\label{fig:PNA_map}Map of the Exclusive
Economic Zones (EEZs) of the region of interest. Countries that belong
to the PNA are shown in blue, while empty polygons indicates all others.
A red line indicates the Kiribati EEZ, and a solid red polygon
delineates PIPA.}
\end{figure}

@@

\begin{figure}
\centering
	\includegraphics{img/all_PS_VDS_cty_year.pdf}
	\caption{\label{fig:VDS_year}a)Vessel-days for all PNA countries (including Tokelau).
	B) Vessel-days for all PNA countries (including Tokelau) by country.}
\end{figure}

\begin{figure}
\centering
	\includegraphics{img/hist_kir_fishing.pdf}
	\caption{\label{fig:hist_kir_fishing}Ridgeplot for the density of the \% of total fishing hours that take place
				within Kiribati EEZ waters by year for treated vessels where the unit of observation is an individual vessel.
}
\end{figure}

@@

@@

%\begin{figure}[h]
%	\includegraphics[width=0.8\textwidth]{img/Color_V1.png}
%	\caption{Map of the Four Protected Areas}
%	\label{fig:maps}
%\end{figure}

@@

\section{References}\label{references}
\bibliography{references}

\hypertarget{appendix}{%
\section{Appendix}\label{appendix}}

>>>>

\documentclass[9pt,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

@@

\begin{abstract}
The Parties to the Nauru Agreement (PNA) is a system for managing the world’s largest skipjack tuna fishery, covering an area of 14.6 million km\textsuperscript{2} in the Pacific Ocean. In 2015, PNA member Kiribati implemented one of the world's largest conservation areas: The Phoenix Island Protected Area (397,447 km\textsuperscript{2}). It is still unclear what the effects of such a big MPA have been and what effects are likely to occur in the future. We use identification of fishing activity via Automatic Identification Systems (AIS) and causal inference techniques evaluate the behavioral changes and spatial redistribution of tuna purse seiners due to the implementation of a Large-Scale Marine Protected Area. Our work provides three main findings: 1) we observe a crowding effect for the first months after implementation, 2) aggregate fishing effort remains relatively constant; and 3) vessels that fished inside the protected area redistribute to adjacent waters, including the high seas. This suggests that rights to fish in PNA waters might now be less valuable. We do not observe an increase in daily fishing hours for vessels that used to fish within the LSMPA, compared to a control group of vessels. Thus, for vessels that have chosen to continue to operate following the closure, we do not observe evidence that they are being forced to exert more effort as a result of the closure. Unfortunately, without detailed data on vessel catch, we cannot conclude whether vessels are better or worse off financially following the closure. We use our results to inform predictions of the impacts of a proposed LSMPA in Palau (a PNA member). Our estimates of the potential losses range from \$2.5 to \$11 million annually. This wide range hinges on what Palau's future PNA allocation will be and the extent to which US aid will continue through non-fishing channels. Our results not only provide an impact evaluation of the effect of LSMPAs on fishing activity, but provide insights into vessel redistribution dynamics, which may have ecological and economic implications for marine conservation and fisheries management. As countries continue to implement LSMPAs as a way to reach ocean protection targets, managers should consider how fishing effort will change in space and time to ensure that fishing effort is not just displaced elsewhere.
\end{abstract}

@@

%(QUESTION FOR JC: WHEN IS THIS (selling of rVDS rights)? And wasn't the closure expected?).

@@

\begin{SCfigure*}
\centering
\includegraphics{img/LSMPAs_map.png}
\caption{\label{fig:LSMPAs_map}Large Scale Marine
Protected Areas. The map shows all areas larger than 250,000
Km\textsuperscript{2}. Areas in dark blue meet at least one of these
conditions: reported no take area is greater than 0, are categorized as
IUCN Ia or Ib, their designated English name is `Protected Area'.}
\end{SCfigure*}

@@

\subsection{Nauru agreement and the Phoenix Islands Protected Area}
\label{nauru-agreement-and-the-phoenix-island-protected-area}

The Nauru Agreement was established in 1982 by a select group of Pacific
island nations to manage their important tuna resources more effectively.
Parties to the Nauru Agreement (PNA) Members
include the Federated States of Micronesia, Kiribati, the Marshall Islands,
Nauru, Palau, Papua New Guinea, the Solomon Islands, and Tuvalu.
Tokelau joined the PNA group in 2012 and started selling access rights in 2013. The Nauru
Agreement regulates access of foreign vessels (\emph{i.e.} those from
non-PNA countries). Holding \textasciitilde{}80\% of historical
skipjack tuna purse seine grounds within their Exclusive Economic Zones, PNA
countries have achieved greater bargaining power when providing fishing access to foreign
fleets \cite{havice_2010}.

The cooperation that emerged under the Nauru Agreement allowed for subsequent
agreements that strengthened fisheries management, like the Palau
Agreement, which limited the number of purse seiners at 205 vessels from
1995-2007.\footnote{See \cite{havice_2010} for a detailed description of the Nauru, Palau, and Federal States of Micronesia agreements.} However, the most notable regulation is
the approach to manage fishing effort: a Vessel Day Scheme (VDS)
implemented in 2007 \citep{havice_2013}. This effectively modified how
fishing effort was managed, from total number of vessels (under the Palau
Agreement) to total vessel-days. 
Under the purse seine VDS, a total number of annual vessel days for the entire fishery is agreed upon by the PNA parties. Each party’s allowable effort (PAE) is then calculated based on historic effort and biomass within each party’s EEZ. Sixty percent of the PAE is calculated based on EEZ effort over the last seven years and 40\% of the PAE is calculated based on the 10-year average of each country’s share of estimated skipjack and yellowfin biomass within its EEZ.\footnote{This is explained in more detail in Article 12.5 of the 2012 Amendment to the Palau Agreement and in \cite{Hagrannsoknir2014}}. A minimum benchmark fee is set for purse seine vessel days which each party can transfer (\emph{i.e.} sell to another PNA member) or sell to the highest bidder (\emph{i.e.} sell to a fishing company). Vessel days can be transferred to fish in any PNA party’s EEZ, without penalty to the transferred parties' allocations \cite{PNA2016}.\footnote{The process of transferring days between parties is described in Article 7 of the Management Scheme \cite{PNA2016}.} The mean value of a vessel day has steadily increased since 2007 \citep{havice_2013}. 
Although detailed records on sales or transfers of vessel days are not publicly available \citep{havice_2013,yeeting2018stabilising}, a minimum benchmark fee of \$8,000/day was set from 2015 \citep{PNA2014a}.

@@

\begin{SCfigure*}
\centering
\includegraphics{img/PNA_map.png}
\caption{\label{fig:PNA_map}Map of the Exclusive
Economic Zones (EEZs) of the region of interest. Countries that belong
to the PNA are shown in blue, while empty polygons indicates all others.
A red line indicates the Kiribati EEZ, and a solid red polygon
delineates PIPA.}
\end{SCfigure*}

@@

\begin{figure}
\centering
	\includegraphics{img/all_PS_VDS_cty_year.pdf}
	\caption{\label{fig:VDS_country_year}A)Observed vessel-days for all VDS countries by fishing and non-fishing days. B) Observed vessel-days for all VDS countries by country.}
\end{figure}

\begin{figure}
\centering
	\includegraphics{img/included_PS_VDS_year_DiD.pdf}
	\caption{\label{fig:included_PS_VDS_year_DiD}Vessel days spent inside VDS waters by group of vessels.}
\end{figure}

\begin{figure}
\centering
	\includegraphics{img/PS_VDS_KIR_by_year.pdf}
	\caption{\label{fig:PS_VDS_KIR_by_year}Vessel days spent inside Kiribati waters by group of vessels.}
\end{figure}

\begin{figure}
\centering
	\includegraphics{img/hist_kir_fishing.pdf}
	\caption{\label{fig:hist_kir_fishing}Ridgeplot for the density of the \% of total fishing hours that take place within Kiribati EEZ waters by year for treated vessels where the unit of observation is an individual vessel.}
\end{figure}

@@

\section{References}\label{references}

\bibliography{references}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%		 					 %%%%%%%%%%%%%
%%%%%%%%%%%%%		 BEGIN APPENDIX		 %%%%%%%%%%%%%
%%%%%%%%%%%%%		 					 %%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Appendix}\label{appendix}

@@

\begin{figure}[h]
	\includegraphics[width=0.8\textwidth]{img/Color_V1.png}
	\caption{Map of the Four Protected Areas}
	\label{fig:maps}
\end{figure}

@@
